---
title: "Practical Machine Learning WriteUp"
author: "zskovesi"
date: "Saturday, October 25, 2014"
output: html_document
---
---

## Background, description of data

The data consists of a set of measurements from sensors attached to subjects performing a variety of physical movements. The task is to prepare a model which will correctly classify activities, given unlabelled sets of observations.

Data can be downloaded from here as we have two datasets:

 - Training: (https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv)
 - Test:(https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv)


#### Process data, data load and cleaning
```{r, echo=TRUE}
# set library and random number generated seed for reproducibility
library(caret)
set.seed(12345)
```

Read training data, manipulate NAs and empty cells, adjust whitespace to have numeric data.

```{r}
Data <- read.csv("pml-training.csv", na.strings=c("NA",""), strip.white=T)
dim(Data)
```

We can exclude time formatted data and NA-s.

```{r}
isNA <- apply(Data, 2, function(x) { sum(is.na(x)) })
targetData <- subset(Data[, which(isNA == 0)], 
                    select=-c(X, user_name, new_window, num_window, raw_timestamp_part_1, raw_timestamp_part_2, cvtd_timestamp))
dim(targetData)
```

We can use the targetData for partitioning and distribution of training and test datasets. For classification we use the *classe* column, pulled from here: (http://groupware.les.inf.puc-rio.br/har)

```{r}
inTrain <- createDataPartition(targetData$classe, p=0.7, list=F)
training <- targetData[inTrain,]
testing <- targetData[-inTrain,]
```

## Training Random Forest Model

The most time consuming part of the report run is modelling Random Forest on the training set. (Using this particular combination of *trControl* parameters is important, as by default bootstrapping is used.)

```{r cache=TRUE, message=FALSE}
ctrl <- trainControl(allowParallel=T, method="cv", number=4)
model <- train(classe ~ ., data=training, model="rf", trControl=ctrl)
pred <- predict(model, newdata=testing)
```

Compare the predictions against the held-back test-set.

```{r}
sum(pred == testing$classe) / length(pred)
confusionMatrix(testing$classe, pred)$table
```

Our forest trained model is 99.2% accurate against our test-set and this is confirmed by the confusion matrix. 

Unknown values are not predicted so we fine tune our model.

```{r fine tuning}
TestData <- read.csv("pml-testing.csv", na.strings=c("NA",""), strip.white=T)
targetTestData <- subset(TestData[, which(isNA == 0)], 
                        select=-c(X, user_name, new_window, num_window, raw_timestamp_part_1, raw_timestamp_part_2, cvtd_timestamp))
predict(model, newdata=targetTestData)
```

List the most significant variables.

```{r}
varImp(model)
```

## Training a filtered Random Forest Model

Filter for the top 10 most significant predictor to train and test a simpler model, that is less time consuming to run.

```{r filtered model, cache=TRUE, message=FALSE}
filtertargetData <- subset(targetData, 
                    select=c(roll_belt, pitch_forearm, yaw_belt, magnet_dumbbell_y, pitch_belt, magnet_dumbbell_z, roll_forearm, accel_dumbbell_y, roll_dumbbell, magnet_dumbbell_x,classe))
filterModel <- train(classe ~ ., data=filtertargetData[inTrain,], model="rf", trControl=ctrl)
```

Its accuracy on the test set is 98.4%.

```{r filtered model accuracy}
predict(filterModel, newdata=targetTestData)
filterPred <- predict(filterModel, newdata=testing)
sum(filterPred == testing$classe) / length(filterPred)
confusionMatrix(testing$classe, filterPred)$table
```

## Training a Support Vector Machine

for accuracy comparison to Random Forest model we run SVM

```{r SVM model, cache=TRUE}
svm <- train(classe ~ ., data=filtertargetData[inTrain,], model="svm", trControl=ctrl)
svmPred <- predict(svm, newdata=testing)
sum(svmPred == testing$classe) / length(svmPred)
confusionMatrix(testing$classe, svmPred)$table
```

It scores 98.3% accuracy and its confusion matrix is only slightly less accurate than the simpler Random Forest model's.

Write answer files for submission
```{r}
answers = rep("pred", 20)
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
```

```{r}
pml_write_files(answers)
```